---
title: "Date Layout"
author: "Yanyu Liang"
date: 2017-03-28
output: html_document
---

<style>
pre code, pre, code {
        white-space: pre !important;
        overflow-x: scroll !important;
        word-break: keep-all !important;
        word-wrap: initial !important;
}
</style>
        
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=2000)
```

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Cell type and annotation type

### Overview

As you can see, some cell types has multiple annotations (either different annotation or different experimental treatments). DNase and CTCF have data from multiple cell types. Also, some TFs, say Pol2, have data sets from the same cell type across multiple treatments.

```{r, fig.height=22, fig.width=20}
library(ggplot2)
library(plyr)
data <- read.table('../data/aucs.txt', sep = '\t', header = T)
data$row_num <- seq(nrow(data))
ggplot(data) + geom_bin2d(aes(x = Cell.Type, y = TF.DNase.HistoneMark)) + theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title=element_text(size=22,face="bold"))
```

### Cell type and annotation type pair that appears multiple times

This may due to multiple experimental treatments but some of them are duplicated experiments. The following lists all duplicated (cell type, annotation type) pair.

```{r}
temp <- aggregate(data[, c('Treatment')], list(data$Cell.Type, data$TF.DNase.HistoneMark), function(x){
        # print(x)
        paste(x, collapse = ' <> ')
})
temp2 <- aggregate(data[, c('Treatment')], list(data$Cell.Type, data$TF.DNase.HistoneMark), function(x){
        # print(x)
        length(x)
})
temp3 <- aggregate(data[, c('row_num')], list(data$Cell.Type, data$TF.DNase.HistoneMark), function(x){
        # print(x)
        paste(x, collapse = ' <> ')
})
temp$num <- temp2$x
temp$row_num <- temp3$x
temp[temp$num > 1,]
```

### Same annotation type, different cell types

Remove all treatments other than "None" (namely ignore complexity from experimental treatment) and count the number of experiments we have per cell type and annotation type pair. Here, CTCF and DNase have many experiments across cell types.

```{r, fig.height=20, fig.width=18}
data_none <- data[data$Treatment == 'None' & !is.na(data$Treatment),]
ggplot(data_none) + geom_bin2d(aes(x = Cell.Type, y = TF.DNase.HistoneMark)) + theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title=element_text(size=22,face="bold"))
```

Here we further explore the performance of DeepSEA on different cell types given a particular annotation type (only consider "None" treatment).

#### Performance distribution within a particular annotation type across cell types

Check out `yanyu_lib.R` [here](#code_r_lib)

```{r}
source('../../cell_type/yanyu_lib.R')
data_none_ctcf <- data_none[data_none$TF.DNase.HistoneMark == 'CTCF',]
data_none_ctcf <- as_num(data_none_ctcf)
p1 <- ggplot(data_none_ctcf) + geom_histogram(aes(x = DeepSEA.ROC.AUC), bins = 20) + ggtitle('ROC AUC in CTCF data sets')
p2 <- ggplot(data_none_ctcf) + geom_histogram(aes(x = DeepSEA.PR.AUC), bins = 20) + ggtitle('PR AUC in CTCF data sets')
multiplot(p1, p2)
data_none_dnase <- data_none[data_none$TF.DNase.HistoneMark == 'DNase',]
data_none_dnase <- as_num(data_none_dnase)
p1 <- ggplot(data_none_dnase) + geom_histogram(aes(x = DeepSEA.ROC.AUC), bins = 20) + ggtitle('ROC AUC in DNase data sets')
p2 <- ggplot(data_none_dnase) + geom_histogram(aes(x = DeepSEA.PR.AUC), bins = 20) + ggtitle('PR AUC in DNase data sets')
multiplot(p1, p2)
```

As you can see, some cell types have low accuracy than others.

## Distance between datasets

For each data set pair, compute how many instances have the same labels and how many have different labels. This setup is done by the following scripts at `/project2/xinhe/yanyul/deep_variant/yanyu/DeepSEA/` and `test_all.h5` is generated from `test.mat` by `/project2/xinhe/yanyul/deep_variant/yanyu/DanQ/my_train/scripts/generate_test_data.py`. The following shows how `/data/test_all.h5.y_disance.hdf5` is generated. Check out `compute_distance.py` [here](#code_dist)

```{bash, eval=F}
$ python my_scripts/compute_distance.py ../DanQ/my_train/data/test_all.h5 my_test/
```

To compute the distance between two sets, Jaccard distance is used here (check definition [here](https://en.wikipedia.org/wiki/Jaccard_index))

```{r}
library(h5)
library(reshape2)
dnase_start <- 1
dnase_end <- 125
tf_start <- 126
tf_end <- 815
hist_start <- 816
hist_end <- 919  # these information is derived from /DeepVariantAnnotation/data/nmeth.3547-S3.xlsx
f <- h5file('../data/test_all.h5.y_disance.hdf5', 'r')
distance <- 1 - f['one_one'][] / (f['one_one'][] + f['zero_one'][] + f['one_zero'][])
melted_distance <- melt(distance)
ggplot(melted_distance) + geom_tile(aes(x = Var1, y = Var2, fill = value)) + scale_fill_gradient(low = "white", high = "black") + geom_vline(xintercept = (dnase_end + tf_start) / 2, color = 'red') + geom_hline(yintercept = (dnase_end + tf_start) / 2, color = 'red') + geom_vline(xintercept = (tf_end + hist_start) / 2, color = 'red') + geom_hline(yintercept = (tf_end + hist_start) / 2, color = 'red') + coord_fixed() + ggtitle('Distance between datasets \n (red line separate DNase/TF/Hist, southwest -> northeast)') + labs(x = 'dataset ID', y = 'dataset ID', fill = 'Jaccard \n distance')
```

The trend is within-group distance is smaller than between-group distance.

### CTCF zoom in

Let's only take a look at instances with CTCF binding (treatment = 'None'). Most of them have some similarity to other members but there are two out liers (ID = 775, 815).

```{r, fig.height=9, fig.width=9}
out <- zoom_in(melted_distance, data_none_ctcf)
sub.ctcf <- out$sub
data_none_ctcf <- out$ori
ggplot() + geom_tile(data = sub.ctcf,aes(x = as.factor(Var1), y = as.factor(Var2), fill = value)) + scale_fill_gradient(low = "white", high = "black") + coord_fixed() + theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text = element_text(size = 5)) + theme(legend.position = "bottom")
```

The following shows the relationship between mean distance within-group and ROC/PR AUC. As you can see, ID 755 and 815 are outliers.

```{r}
p1 <- ggplot(data_none_ctcf, aes(x = dist.mean, y = DeepSEA.ROC.AUC, label = row_num)) + geom_text()
p2 <- ggplot(data_none_ctcf, aes(x = dist.mean, y = DeepSEA.PR.AUC, label = row_num)) + geom_text()
multiplot(p1, p2, cols=1)
```

For your interest, the following shows the cell types that either have worst ROC/PR AUC or are far from others.

Select the ones which are furthest to others.

```{r}
data[data_none_ctcf[order(data_none_ctcf$dist.mean, decreasing = T)[1:5],]$row_num,]
```

Select the ones which gets lowest ROC AUC scores.

```{r}
data[data_none_ctcf[order(data_none_ctcf$DeepSEA.ROC.AUC)[1:5],]$row_num,]
```

Select the ones which gets lowest PR AUC scores.

```{r}
data[data_none_ctcf[order(data_none_ctcf$DeepSEA.PR.AUC)[1:5],]$row_num,]
```

### DNase zoom in

The same analysis is done on DNase group (treatment = 'None').

```{r, fig.height=9, fig.width=9}
out <- zoom_in(melted_distance, data_none_dnase)
sub.dnase <- out$sub
data_none_dnase <- out$ori
ggplot() + geom_tile(data = sub.dnase,aes(x = as.factor(Var1), y = as.factor(Var2), fill = value)) + scale_fill_gradient(low = "white", high = "black") + coord_fixed() + theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text = element_text(size = 5)) + theme(legend.position = "bottom")
```
```{r}
p1 <- ggplot(data_none_dnase, aes(x = dist.mean, y = DeepSEA.ROC.AUC, label = row_num)) + geom_text()
p2 <- ggplot(data_none_dnase, aes(x = dist.mean, y = DeepSEA.PR.AUC, label = row_num)) + geom_text()
multiplot(p1, p2, cols=1)
```

Select the ones which are furthest to others.

```{r}
data[data_none_dnase[order(data_none_dnase$dist.mean, decreasing = T)[1:5],]$row_num,]
```

Select the ones which gets lowest ROC AUC scores.

```{r}
data[data_none_dnase[order(data_none_dnase$DeepSEA.ROC.AUC)[1:5],]$row_num,]
```

Select the ones which gets lowest PR AUC scores.

```{r}
data[data_none_dnase[order(data_none_dnase$DeepSEA.PR.AUC)[1:5],]$row_num,]
```


## Codes 

### `yanyu_lib.R` {#code_r_lib}

```{bash, echo=F}
cat ../../cell_type/yanyu_lib.R
```

### `compute_distance.py` {#code_dist}

```{bash, echo=F}
cat ../../../my_scripts/compute_distance.py 
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```


