---
title: "SVD of DeepSea response"
author: 
- "Nicholas Knoblauch"
- Yanyu Liang
date: 2017-03-28
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```




<!-- Add your analysis here -->

This notebook shows the SVD of the DeepSea response matrix.

```{r}
library(ggplot2)
```


The mean was subtracted from each column before singular values were computed.


## Loading data
```{r}
evals_df <- readRDS("../data/DeepSea_evals_df.RDS")
deepsea_pca_df <- readRDS("../data/DeepSeaPCA_df.RDS")
```



```{r}
ggplot(evals_df,aes(x=ind,y=evals/sum(evals)))+geom_point()+ggtitle("Scree plot")+xlab("Rank")+ylab("Eigenvalues")
```
The Scree plot shows us that the first two eigenvalues (PC's) contain a significant amount of the variance, but that there's no sharp dropoff.


## PCA plots

```{r}
ggplot(deepsea_pca_df,aes(x=PC_1,y=PC_2,col=Source))+geom_point()+ggtitle("PC 1  1 vs PC 2")
```

Plot by annotation types (TF Binding/DNase/Histone Mark). See `to_annotation_type` function [here](#code)
```{r}
library(stringr)
source('../../cell_type/yanyu_lib.R')
aucs_danq <- read.table('../data/aucs.with.tissue.txt', sep = '\t', header = T, fill = T)
aucs_danq$DeepSEA.ROC.AUC <- as.numeric(as.character(aucs_danq$DeepSEA.ROC.AUC))
aucs_danq <- aucs_danq[!is.na(aucs_danq$DeepSEA.ROC.AUC),]
deepsea_pca_df$AnnotationType <- to_annotation_type(aucs_danq$TF.DNase.HistoneMark)
deepsea_pca_df$ROC.AUC <- aucs_danq$DeepSEA.ROC.AUC
deepsea_pca_df$PR.AUC <- as.numeric(as.character(aucs_danq$DeepSEA.PR.AUC))
cleaned_tissue <- as.character(aucs_danq$Tissue)
cleaned_tissue[!cleaned_tissue %in% names(table(cleaned_tissue))[table(cleaned_tissue) > 30]] <- 'Minor'
deepsea_pca_df$Tissue <- cleaned_tissue
cleaned_karyotype <- as.character(aucs_danq$Karyotype)
cleaned_karyotype[cleaned_karyotype == ''] <- 'unknown'
deepsea_pca_df$Karyotype <- cleaned_karyotype
ggplot(deepsea_pca_df,aes(x=PC_1,y=PC_2,col=AnnotationType))+geom_point()+ggtitle("PC 1  1 vs PC 2")
```

PC1 and PC2 separate Histone marks, TF binding, and DNase.

```{r}
ggplot(deepsea_pca_df,aes(x=PC_1,y=PC_2,col=Tissue))+geom_point()+ggtitle("PC 1  1 vs PC 2")
```

```{r}
ggplot(deepsea_pca_df,aes(x=PC_1,y=PC_2,col=Karyotype))+geom_point()+ggtitle("PC 1  1 vs PC 2")
```

## Code {#code}

```{r}
to_annotation_type
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
