---
title: "Computing PCA from SVD"
author: "Nicholas Knoblauch"
date: 2017-03-28
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r}
library(readr)
library(h5)
library(dplyr)
```


To compute the PCA of the response matrix, we first scaled the columns of the matrix (see IPython notebook) and then computed singular values.  As a refresher, remember that for a matrix $X$ (for which the mean of each column is $0$), the covariance matrix $C$ is
$$C= X^{T}X/(n-1)$$ 

The eigenvalue decomposition of the covariance matrix C is gives us the principle components:

$$C=VLV^{T}$$

Remember that for any matrix $X$, the Singular Value Decomposition of that matrix $X$  is 

$$X = USV^{T}$$

It's easy to show that
$$C=VSU^{T}USV^{T}/(n-1)= V\frac{S^2}{n-1}V^{T}$$ 

This means that $US$ are the principle components of $X$


We'll first pull the features from the DeepSEA website
```{r}
feature_url <- "http://deepsea.princeton.edu/media/help/features.txt"
features <- read_delim(feature_url,delim="\t",col_names = c("Source",
                                                            "CellType",
                                                            "DataType",
                                                            "Treatment",
                                                            "AUC"),skip = 1)
```

Now we'll load the svd we computed in `dask`

```{r}

train_svdf <- "/media/nwknoblauch/Data/DeepSEA/train_svd_50_3.h5"

tsvdf <- h5file(train_svdf,mode='r')

tu <- tsvdf["U"][]
td <- tsvdf["D"][]
tv <- tsvdf["V"][]
s_tu <- tu*td
```



Save the results

```{r}
tu_df <- as_data_frame(s_tu)
colnames(tu_df) <- paste0("PC_",1:50)

d_df <- data_frame(evals=td^2/(ncol(tv)-1),ind=1:length(td))
#saveRDS(d_df,"../data/DeepSea_evals_df.RDS")

features <- mutate(features,AUC=as.numeric(AUC)) %>% filter(!is.na(AUC))
ntu_df <- bind_cols(features,slice(tu_df,-1))
#saveRDS(ntu_df,"../data/DeepSeaPCA_df.RDS")
```








## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
